The reference model approach in feature selection problems \vspace{.1in}

### 2019-06-13 paper/ref_approach.tex
updated version

%In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, which can be seen as a special case of model reduction. From a full Bayesian perspective, when the goal of the analysis is to do prediction, no selection should be carried out and integration over all parameters uncertainties, i.e. posterior distributions, is the recommended procedure. In a high dimensional parameter space the choice of the prior plays an even more key role due to different difficulties that can rise, as computation burden and overfitting. In addition to that, sometimes collecting all the features for a future observation is not possible or too expensive. In order to solve or, at least, simplify these problems, different feature selection procedures have been proposed both in the frequentist and the Bayesian setting \cite[e.g. see][]{paper:feature_selection,paper:vehtari_ojanen,paper:model_selection}. Sometimes the task of the analysis is not prediction, but rather inference on which variables are substantially related to the phenomenon of interest. This typically happens in a \textit{small n large p} scenario, as in many microarray data analysis in medicine or biology. A very large number of genes is usually analysed in two groups of individuals labeled as \textit{positive} or \textit{negative} to some disease. The main goal of the analysis is to spot the subset of explanatory genes in order to perform further experiments \cite[see examples in][]{book:efron}.

In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, which is a special case of model reduction. The necessity of carrying out the selection arises from different problematics. Sometimes the leading objective of the analysis is to predict future observations, thus the main focus in building the model is on the predictive performance. Yet a high dimensional parameter space often leads to dangers as overfitting or other technical problems, as expensive costs to collect all the features or computational and memory burdens (e.g. when we are operating on limited resources), and thus a sparser representation of the model would be preferred. In addition to that, simpler models bring also advantages in terms of interpretability. Some other times, variable selection can be also seen as an inferential tool: the interest is on knowing which variables are meaningful, or relevant, in relation to the target variable of interest. This typically happens in a \textit{small n large p} scenario, as in many microarray data analysis in medicine or biology. A very large number of genes is usually analysed in two groups of individuals labeled as \textit{positive} or \textit{negative} to some disease; the goal is to spot the subset of explanatory genes in order to perform further experiments \cite[see examples in][]{book:efron}.

Either the target of the selection is the minimal subset of feature with good enough predictive ability, or the whole subset of relevant variables, the selection process is going to be based on assumptions regarding the ``true'' data generation mechanism. Making assumptions means making models, thus we identify two families of feature selection, or more in general model reduction, approaches depending on how the data generation mechanism is modelled: the data based and the reference model based ones. The former includes all those methods that use the observed empirical distribution as model approximation of the data generation mechanism, as for example happens in the Lasso selection \cite{paper:tibshirani_lasso} or in the stepwise backward/forward regression, where the selected submodels balance sparsity with predictive ability on the observed data. The reference model based methods use instead the predictive distribution of a full-encompassing model, i.e. the reference model, as an approximation of the data generation mechanism. Such an idea is widely present in the literature with different names; for example, \cite{book:harrell} refers to the reference model as a full model that can be thought as the ``gold standard'' and shows some examples of how it can be used to look for a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model. \cite{paper:faraggi_nn} deal with the necessity of identifying interpretable risk groups in the context of survival data using neural networks, which typically perform very well in terms of prediction, but whose covariates are difficult to be understood in terms of relevance. \cite{paper:faraggi_nn} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups. \cite{paper:paul_preconditioning}, using the term preconditioning, explore approximating models fitting Lasso or stepwise regression against consistent estimates $\hat{y}$ of a reference model instead of the observed response quantities $y$. 
%\cite{book:harrell} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.

@@

The reference model approach has been used for long time also in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. In the case of generalised linear models (GLMs), \cite{paper:tran_predictivelasso} provide a procedure to estimate and do variable selection with the goal of prediction: the best approximating model is recovered by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. The resulting problem is convex and thus many efficient optimisation algorithm are available. Following \cite{paper:goutis_projection}, \cite{paper:original_proj}  instead project the posterior distribution of the reference model in each candidate model and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models being GLMs from the exponential family, the optimisation problem corresponds to the task of finding the maximum likelihood estimate using the expected pointwise predictions of the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{paper:nott_projection} that add an $l_{1}$ penalty to select the best approximating model. \cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance of the explored submodels with the reference model one.

@@

The paper is structured as follows. In Section \ref{reference-model-approach} we introduce the idea of the reference model, its benefits with examples, including the projection predictive approach, and how it can be used as a filter on data in a simple way. In Section \ref{comparison} we use different methods for feature selection comparing the selection results using or not the reference model in the framework of the normal means problem. Finally, in Section \ref{discussion} we present our conclusions.

@@

The main assumption under any reference model approach is to be operating in a $\mathcal{M}$-\textit{complete} framework \citep{book:bernardo_smith,paper:vehtari_ojanen}, i.e. to be able to construct a model, that is the reference model, whose description of a future observation distribution is comparable to the ``true'' data generation mechanism. In practice, any full-encompassing model which has a good predictive performance is a candidate reference model. General modelling guidelines apply also for the reference model, except no variable selection is carried out at this stage.

A good predictive model is able to clean part of the noise present in the data. The noise is the main source of the instability and tends to obscure the relevance of the covariates in relation to the target variable of interest. Indeed, consider the following data generation mechanism for each statistical unit, suggested in \cite{paper:projpred}:
\
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber
\end{alignat}
$f$ is the latent variable of interest of which $Y$ is a noisy observation. The first $k$ covariates are significantly related to the target variable $Y$ and correlated among themselves. Precisely $\rho$ is the correlation among any pair of the first $k$ covariates, whereas $\sqrt{\rho}$ and $\sqrt{\rho}/2$ are the level of correlation between any significant covariate and respectively $f$ and $Y$. If we had an infinite amount of observations, the sample correlation would be equal to the true correlation, though, even in this ideal asymptotic regime, the correlation computed would be still a biased indicator of the true relevance of each variable with respect to the latent variable of interest due to the intrinsic noisy nature of $Y$. When using a reference model, we are trying to model $f$ through the covariates $\{X_{j}\}_{j=1}^{p}$ taking in account that what we observe (i.e. $Y$) is corrupted by noise. Thus, if our model is good, we are able to describe $f$ better than the simply observed variable $Y$ and to use such information to assess the relevance of the features. Indeed, Figure \ref{fig:correlation} shows the scatter plot of the absolute value of the sample correlation of each feature with the point prediction by a reference model, in this case see model \eqref{eq:ref_mod}, against the absolute value of the sample correlation of each feature with the noisy observations $\{y\}_{i=1}^{n}$. Looking at the marginal distribution, we observe that using a reference model the two group of features, i.e. relevant and non-relevant, have a minor overlap, and thus can be better distinguished, with respect to when the correlation is computed using directly the observed data. We can think about the correlation as a general indicator of the relevance of a feature.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.3\textwidth]{graphics/correlation.pdf}
  \caption{Sample correlation plot of each feature (relevant in red, non-relevant in blue) with the target variable $y$ and the latent variable $f$ respectively on the x and y axis. Simulation parameters: $n=70$, $\rho=0.3$, $k=100$, $p=1000$.\\}
  \label{fig:correlation}
\end{figure}

In Figure \ref{fig:correlation} the reference model acts as a noise-filter on the observed data through its predictions. This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paper:paul_preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
As an another example, let us consider a candidate submodel in a feature selection process and let us refer with $\pi$ to the parameter distribution of the submodel and with $q_{\pi}(\tilde{y})$ to the induced predictive distribution. We would like to choose $\pi$ in order that the latter maximises some predictive performance utility, as the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
\end{equation}
If we refer to the posterior predictive distribution of a reference model with $p_{\text{ref}}(\tilde{y}|D)$, where $D$ stands for the data on which we condition on, we can approximate \eqref{eq:elpd} using $p_{\text{ref}}(\tilde{y}|D)$ instead of the true, unknown, data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the KL divergence between the reference model predictive distribution and the submodel one:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred as the projection of the predictive distribution, which is the general idea behind the projection predictive approach. Again the reference model is acting as a filter, or substitute, on data.

@@

%What we do in Figure \ref{fig:correlation} is to use the reference model predictive ability as a noise-filter on the observed data. This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paper:paul_preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}. Also the general idea of the projection of the predictive can be seen under this perspective. Indeed, let us call $p_{\text{ref}}(\tilde{y}|D)$ the posterior predictive distribution for a new observation $\tilde{y}$ for the reference model and let $\pi(\theta)$ and $q_{\pi}(\tilde{y})$ be respectively the submodel parameter and the induced predictive distributions. The projection of the predictive distribution allows to find the optimal parameter distribution $\pi^{*}$ minimising the KL discrepancy among the induced predictive distribution of the submodel and the reference model one:
%\
%\begin{equation} 
%\pi^{*}\quad \text{s.t.} \quad \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi^{*}}(\tilde{y})]=\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
%\end{equation}
%Consider now the predictive performance of such submodel, a common utility in Bayesian context is the expected log-predictive density defined as:
%\
%\begin{equation} 
%\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
%\end{equation}
%where $p_{t}$ is the distribution of a new observation under the true data generation mechanism. Maximising the elpd of a model is equivalent to minimising the KL-divergence of its predictive distribution from the true data generation mechanism. Therefore the projection corresponds to maximising the elpd using as distribution of the true data generation mechanism the posterior predictive distribution of the reference model, i.e.:
%\
%\begin{equation} \label{proj_as_filter}
%\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] \leftrightarrow \underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} 
%\end{equation}
%Equation \eqref{proj_as_filter} clearly shows the action of the reference model as a filter on our target observations, substituting the unknown true data generation mechanism, that is usually estimated with the data empirical distribution, with the best known approximation that is available, i.e. the reference model posterior predictive distribution.

@@

The projection predictive approach uses the reference model to project the posterior samples in the submodels. The R-package \textit{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and additionally provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. When the goal is prediction, the choice is usually on the smallest submodel that is sufficiently close to the reference model prediction utility score. Further details on the projection predictive approach can be found in \cite{paper:projpred}.

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{paper:bodyfat_johnson}. The analysis is mainly taken from Vehtari's R-notebook available at \url{https://avehtari.github.io/modelselection/bodyfat.html}, adjusting only the number of bootstrap repetitions to 1000 to have a consistent comparison with results from \cite{paper:bodyfat}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure, and the original covariates are 13 anthropometric measurements (e.g. height, weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection, for a total of 251 observations. The goal is to find the simplest model which is able to describe, that is to predict, sufficiently well the amount of fat.

@@

We can consider the reference model approach as a family of methods, thus providing an answer to the question what the computational burden is, is not possible, because not unique. It is possible to identify a common and inevitable overhead though, that is the cost of fitting the reference model. In the simple reference model approach we have described at the end of Section \ref{reference-model-approach}, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance, however including all the available features can be often quite computationally demanding. In such a case a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{paper:original_spc,paper:piironen_spc}. However, in some cases this does not always result in an appreciable speed-up, as when there are many predictive and not correlated features and, thus, many of them have to be included in the reference model anyway. Other alternatives could be fixing some parameters with their maximum marginal likelihood value (typically referred as empirical Bayes) or even use a frequentist approach, if predictions are good enough.

@@

Note that it is equivalent to a linear regression where the design matrix is the identity one. This formulation can be retrieved in different ways, a common example is given by microarray data, where a large set of genes are usually tested in two groups of patient labelled as positive or negative to some disease. The objective of the analysis is to spot the whole subset of statistically relevant genes to the disease and one possible way to proceed is to compute the two-samples $t$-statistics for each gene and after combining it with the cumulative density function of the standard distribution, the resulting data can be used in the problem formulation \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, book:efron}. 

@@

In our examples we retrieve the normal means problem from the sample correlation using the Fisher $z$-transformation approximation \citep{paper:hawkins}. Suppose to have a continuous target random variable $Y$ and a set of $p$ continuous covariates $\{X_{j}\}_{j=1}^{p}$ and let us call $\rho_{j}=Cor(Y,X_{j})$. Suppose to observe $n$ statistical units and define $r_{j}$ the sample correlation between the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the j-th covariate $\{x_{ij}\}_{i=1}^{n}$. Let finally refer to the function $\text{tanh}^{-1}(\cdot)$ as $T_{F}(\cdot)$. It holds the following modelling approximation (assuming each pair $(Y,X_{j})$ normally distributed):

@@

where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value, $\hat{\theta}_{j}$ is the posterior median of each parameter $\theta_{j}$ and \eqref{eq:SESE} is a sum of posterior expactations. We divide the errors by $(n-3)$ to have quantities independent from the number of statistical units. Note that the quantity SESE takes in account also the uncertainty, thus the shrinkage, of the posterior distributions, whereas the SSE measures only the bias in the point estimates.

>>>>

%In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, which can be seen as a special case of model reduction. From a full Bayesian perspective, when the goal of the analysis is to do prediction, no selection should be carried out and integration over all parameters uncertainties, i.e. posterior distributions, is the recommended procedure. In a high dimensional parameter space the choice of the prior plays an even more key role due to different difficulties that can rise, as computation burden and overfitting. In addition to that, sometimes collecting all the features for a future observation is not possible or too expensive. In order to solve or, at least, simplify these problems, different feature selection procedures have been proposed both in the frequentist and the Bayesian setting \cite[e.g. see][]{paper:feature_selection,vehtari2012survey,paper:model_selection}. Sometimes the task of the analysis is not prediction, but rather inference on which variables are substantially related to the phenomenon of interest. This typically happens in a \textit{small n large p} scenario, as in many microarray data analysis in medicine or biology. A very large number of genes is usually analysed in two groups of individuals labeled as \textit{positive} or \textit{negative} to some disease. The main goal of the analysis is to spot the subset of explanatory genes in order to perform further experiments \cite[see examples in][]{efron2012large}.

In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, which is a special case of model reduction. The necessity of carrying out the selection arises from different problematics. Sometimes the leading objective of the analysis is to predict future observations, thus the main focus in building the model is on the predictive performance. Yet a high dimensional parameter space often leads to dangers as overfitting or other technical problems, as expensive costs to collect all the features or computational and memory burdens (e.g. when we are operating on limited resources), and thus a sparser representation of the model would be preferred. In addition to that, simpler models bring also advantages in terms of interpretability. Some other times, variable selection can be also seen as an inferential tool: the interest is on knowing which variables are meaningful, or relevant, in relation to the target variable of interest. This typically happens in a \textit{small n large p} scenario, as in many microarray data analysis in medicine or biology. A very large number of genes is usually analysed in two groups of individuals labeled as \textit{positive} or \textit{negative} to some disease; the goal is to spot the subset of explanatory genes in order to perform further experiments \cite[see examples in][]{efron2012large}.

Either the target of the selection is the minimal subset of feature with good enough predictive ability, or the whole subset of relevant variables, the selection process is going to be based on assumptions regarding the ``true'' data generation mechanism. Making assumptions means making models, thus we identify two families of feature selection, or more in general model reduction, approaches depending on how the data generation mechanism is modelled: the data based and the reference model based ones. The former includes all those methods that use the observed empirical distribution as model approximation of the data generation mechanism, as for example happens in the Lasso selection \cite{tibshirani1996regression} or in the stepwise backward/forward regression, where the selected submodels balance sparsity with predictive ability on the observed data. The reference model based methods use instead the predictive distribution of a full-encompassing model, i.e. the reference model, as an approximation of the data generation mechanism. Such an idea is widely present in the literature with different names; for example, \cite{harrell2015regression} refers to the reference model as a full model that can be thought as the ``gold standard'' and shows some examples of how it can be used to look for a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model. \cite{faraggi2001understanding} deal with the necessity of identifying interpretable risk groups in the context of survival data using neural networks, which typically perform very well in terms of prediction, but whose covariates are difficult to be understood in terms of relevance. \cite{faraggi2001understanding} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups. \cite{paul2008preconditioning}, using the term preconditioning, explore approximating models fitting Lasso or stepwise regression against consistent estimates $\hat{y}$ of a reference model instead of the observed response quantities $y$. 
%\cite{harrell2015regression} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.

@@

The reference model approach has been used for long time also in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. In the case of generalised linear models (GLMs), \cite{tran2012predictive} provide a procedure to estimate and do variable selection with the goal of prediction: the best approximating model is recovered by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. The resulting problem is convex and thus many efficient optimisation algorithm are available. Following \cite{goutis1998model}, \cite{paper:original_proj}  instead project the posterior distribution of the reference model in each candidate model and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models being GLMs from the exponential family, the optimisation problem corresponds to the task of finding the maximum likelihood estimate using the expected pointwise predictions of the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{nott2010bayesian} that add an $l_{1}$ penalty to select the best approximating model. A recent review of these methods together with some methodological improvements is given by \cite{paper:projpred}.
%\cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance of the explored submodels with the reference model one.

@@

The paper is structured as follows. In Section \ref{reference-model-approach} we introduce the idea of the reference model, its benefits with examples, including the projection predictive approach, and how it can be used as a filter on data in a simple way. In Section \ref{comparison} we use different methods for feature selection comparing the selection results using or not the reference model in the framework of the normal means problem. Finally, in Section \ref{conclusion} we present our conclusions.

@@

The main assumption under any reference model approach is to be operating in a $\mathcal{M}$-\textit{complete} framework \citep{book:bernardo_smith,vehtari2012survey}, i.e. to be able to construct a model, that is the reference model, whose description of a future observation distribution is comparable to the ``true'' data generation mechanism. In practice, any full-encompassing model which has a good predictive performance is a candidate reference model. General modelling guidelines apply also for the reference model, except no variable selection is carried out at this stage.

A good predictive model is able to clean part of the noise present in the data. The noise is the main source of the instability and tends to obscure the relevance of the covariates in relation to the target variable of interest. Indeed, let us consider the following explanatory example taken from \cite{paper:projpred}. The data generation mechanism for each statistical unit is the following:
\
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber
\end{alignat}
$f$ is the latent variable of interest of which $Y$ is a noisy observation. The first $k$ covariates are significantly related to the target variable $Y$ and correlated among themselves. Precisely, $\rho$ is the correlation among any pair of the first $k$ covariates, whereas $\sqrt{\rho}$ and $\sqrt{\rho}/2$ are the level of correlation between any significant covariate and respectively $f$ and $Y$. If we had an infinite amount of observations, the sample correlation would be equal to the true correlation, though, even in this ideal asymptotic regime, the correlation computed would be still a biased indicator of the true relevance of each variable with respect to the latent variable of interest due to the intrinsic noisy nature of $Y$. When using a reference model, we are trying to model $f$ through the covariates $\{X_{j}\}_{j=1}^{p}$ taking in account that what we observe (i.e. $Y$) is corrupted by noise. Thus, if our model is good, we are able to describe $f$ better than the simply observed variable $Y$ and to use such information to assess the relevance of the features. Indeed, Figure \ref{fig:correlation} shows the scatter plot of the absolute value of the sample correlation of each feature with the point prediction by a reference model, in this case see model \eqref{eq:ref_mod}, against the absolute value of the sample correlation of each feature with the noisy observations $\{y\}_{i=1}^{n}$. Looking at the marginal distribution, we observe that using a reference model the two group of features, i.e. relevant and non-relevant, have a minor overlap, and thus can be better distinguished, with respect to when the correlation is computed using directly the observed data. We can think about the correlation as a general indicator of the relevance of a feature.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.3\textwidth]{graphics/correlation.pdf}
  \caption{Sample correlation plot of each feature (relevant in red, non-relevant in blue) with the target variable $y$ and the latent variable $f$ respectively on the x and y axis. Data are generated according to procedure \eqref{eq:simulated_data} with parameters $n=70$, $\rho=0.3$, $k=100$, $p=1000$.\\}
  \label{fig:correlation}
\end{figure}

In Figure \ref{fig:correlation} the reference model acts as a noise-filter on the observed data through its predictions. This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
As an another example, let us consider a candidate submodel in a feature selection process and let us refer with $\pi$ to the parameter distribution of the submodel and with $q_{\pi}(\tilde{y})$ to the induced predictive distribution. We would like to choose $\pi$ in order that the latter maximises some predictive performance utility, as the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
\end{equation}
If we refer to the posterior predictive distribution of a reference model with $p_{\text{ref}}(\tilde{y}|D)$, where $D$ stands for the data on which we condition on, we can approximate \eqref{eq:elpd} using $p_{\text{ref}}(\tilde{y}|D)$ instead of the true, unknown, data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the KL divergence between the reference model predictive distribution and the submodel one:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred as the projection of the predictive distribution, which is the general idea behind the projection predictive approach. Again the reference model is acting as a filter, or substitute, on data.

@@

%What we do in Figure \ref{fig:correlation} is to use the reference model predictive ability as a noise-filter on the observed data. This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}. Also the general idea of the projection of the predictive can be seen under this perspective. Indeed, let us call $p_{\text{ref}}(\tilde{y}|D)$ the posterior predictive distribution for a new observation $\tilde{y}$ for the reference model and let $\pi(\theta)$ and $q_{\pi}(\tilde{y})$ be respectively the submodel parameter and the induced predictive distributions. The projection of the predictive distribution allows to find the optimal parameter distribution $\pi^{*}$ minimising the KL discrepancy among the induced predictive distribution of the submodel and the reference model one:
%\
%\begin{equation} 
%\pi^{*}\quad \text{s.t.} \quad \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi^{*}}(\tilde{y})]=\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
%\end{equation}
%Consider now the predictive performance of such submodel, a common utility in Bayesian context is the expected log-predictive density defined as:
%\
%\begin{equation} 
%\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
%\end{equation}
%where $p_{t}$ is the distribution of a new observation under the true data generation mechanism. Maximising the elpd of a model is equivalent to minimising the KL-divergence of its predictive distribution from the true data generation mechanism. Therefore the projection corresponds to maximising the elpd using as distribution of the true data generation mechanism the posterior predictive distribution of the reference model, i.e.:
%\
%\begin{equation} \label{proj_as_filter}
%\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] \leftrightarrow \underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} 
%\end{equation}
%Equation \eqref{proj_as_filter} clearly shows the action of the reference model as a filter on our target observations, substituting the unknown true data generation mechanism, that is usually estimated with the data empirical distribution, with the best known approximation that is available, i.e. the reference model posterior predictive distribution.

@@

As a motivational example to the study of the consequences of the use of a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples in the submodels. The R-package \textit{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Assume to have $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates, the main step are the followings:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of S posterior samples of the vector parameter of the model.
\item Rank the covariates using a heuristics and consider as candidate submodels the nested ones, starting to include the best ranked feature. The submodel are naturally identify by their model size.
\item For each submodel, project the reference model samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL}\;[p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\,||\,p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})] \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the model distribution of the reference model with parameters fixed at the sample value $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariates related to the statistical unit (identified by the subscript $i$ in the variable $\tilde{y}_{i}$), whereas $p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ the same but for the submodel.
\item For each submodel (size), test the predictive performance.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model prediction utility score.
\end{enumerate}

Expression \eqref{eq:draw_by_draw} is not a easy an easy optimisation problem, indeed \textit{projpred} implements the projection in the special case of submodels in the GLMs family, where \eqref{eq:draw_by_draw} reduces to a maximum likelihood estimation problem. For further details on \textit{projpred} workflow and implementation see \cite{paper:projpred}.

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{johnson1996fitting}. The analysis is mainly taken from Vehtari's R-notebook available at \url{https://avehtari.github.io/modelselection/bodyfat.html}, adjusting only the number of bootstrap repetitions to 1000 to have a consistent comparison with results from \cite{paper:bodyfat}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure, and the original covariates are 13 anthropometric measurements (e.g. height, weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection, for a total of 251 observations. The goal is to find the simplest model which is able to describe, that is to predict, sufficiently well the amount of fat.

@@

We can consider the reference model approach as a family of methods, thus providing an answer to the question what the computational burden is, is not possible, because not unique. It is possible to identify a common and inevitable overhead though, that is the cost of fitting the reference model. In the simple reference model approach we have described at the end of Section \ref{reference-model-approach}, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance, however including all the available features can be often quite computationally demanding. In such a case a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases this does not always result in an appreciable speed-up, as when there are many predictive and not correlated features and, thus, many of them have to be included in the reference model anyway. Other alternatives could be fixing some parameters with their maximum marginal likelihood value (typically referred as empirical Bayes) or even use a frequentist approach, if predictions are good enough.

@@

Note that it is equivalent to a linear regression where the design matrix is the identity one. This formulation can be retrieved in different ways, a common example is given by microarray data, where a large set of genes are usually tested in two groups of patient labelled as positive or negative to some disease. The objective of the analysis is to spot the whole subset of statistically relevant genes to the disease and one possible way to proceed is to compute the two-samples $t$-statistics for each gene and after combining it with the cumulative density function of the standard distribution, the resulting data can be used in the problem formulation \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, efron2012large}. 

@@

In our examples we retrieve the normal means problem from the sample correlation using the Fisher $z$-transformation approximation \citep{hawkins1989using}. Suppose to have a continuous target random variable $Y$ and a set of $p$ continuous covariates $\{X_{j}\}_{j=1}^{p}$ and let us call $\rho_{j}=Cor(Y,X_{j})$. Suppose to observe $n$ statistical units and define $r_{j}$ the sample correlation between the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the j-th covariate $\{x_{ij}\}_{i=1}^{n}$. Let finally refer to the function $\text{tanh}^{-1}(\cdot)$ as $T_{F}(\cdot)$. It holds the following modelling approximation (assuming each pair $(Y,X_{j})$ normally distributed):

@@

where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value, $\hat{\theta}_{j}$ is the posterior median of each parameter $\theta_{j}$ and \eqref{eq:SESE} is a sum of posterior expectations. We divide the errors by $(n-3)$ to have quantities independent from the number of statistical units. Note that the quantity SESE takes in account also the uncertainty, thus the shrinkage, of the posterior distributions, whereas the SSE measures only the bias in the point estimates. Note that any kind of point estimate could be used to compute the SSE, here we follow \cite{paper:dirichlet_laplace} using the posterior median. The results that we show later in Section \ref{shrinkage-signal} are not sensible to such a choice. We also take into account the posterior mean for the shrinkage factor ($\kappa_{j}$) for each parameter $\theta_{j}$, which for model \eqref{eq:normal_means_problem2} is defined as follows:
\begin{equation}
\kappa_{j}^{\text{RHS}}=\frac{1}{1+\tau^{2}\lambda_{j}^{2}}, \qquad \kappa_{j}^{\text{DL}}=\frac{1}{1+\psi\phi_{j}^{2}\tau^{2}}
